---
title: "MonopolyWork"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. _Rules!_ Monopoly has several rules related to movement around the game board. Movement is affected by standard dice rolls, special conditions for some dice rolls, and game cards. Find and summarize these rules. 

SET UP: Each player is given $3,00 and starts on GO (position 1).
GAME PLAY: Player one goes first followed by player 2, player 3, and then player 4. Each player rolls the dice on their turn and follows the instructions for the place they land on.
ROLLING DOUBLES: Players who roll doubles must roll again. Players move their piece and follow instructions of the space they land on between rolls. 3 sets of doubles send the player to jail. 
SPECIAL SPACES:
GO: Every time players pass go they collect $50.
PROPERTIES: Players must buy a property if they have enough money to. If a player lands on a property another player owns they must pay the specified rent amount.
JAIL: Players who get sent to jail must pay the $50 fee to begin their next turn and their position is reset to jail.
GAME END: The game ends when only one player remains. Each player gets eliminated when the money in their account reaches $0 or below.

SPECIAL RULES: Houses and hotels have been removed. Likewise so have hotel and house related chance and community chest cards. Get out of jail free cards have been removed for simplicity as well. Players cannot roll doubles to get out of jail, they must pay the fee. Auctions have been removed as well as selling/mortgaging properties.
2. _Let's Get Moving_ Build a system that allows for up to 4 players to move around the game board. As a first attempt, focus on standard dice rolls, ignoring additional rules related to rolling doubles, jail, or special cards. Which properties are landed on most frequently? Create a visualization for this!
3. _Double Down_ Improve your system to incorporate going to jail, including the doubles-rolling condition. Which properties are landed on most frequently now? Has this changed dramatically from the previous system (that didn't account for jail)? Again, visualize this!

4. _Take a Chance_ Incorporate the movement cards into your system. You can ignore the effects of non-movement cards, but not their existence! Which properties are landed on most frequently now? Has this changed dramatically from the previous system (that didn't account for Chance)? Again, visualize this!

5. _Play a Game_ Extend your system to allow for up to 4 players to "play" the game. You may choose a reasonable starting board and player money level (so, you don't have to buy properties). Include payments between players, payments to the bank, and payments from the bank (there are several cards that will effect these that you should account for). You do not need to consider mortgaging properties (so a player loses when their liquid cash has run out). How long do games typically last given your starting board and money level? How does this change if money paid to the bank is placed in "Free Parking" and awarded to players landing on that space?

```{r}
#creating the board
tiles <- c("GO", "Mediterranean Avenue", "Community Chest", "Baltic Avenue", "Income Tax", "Reading Railroad", "Oriental Avenue","Chance","Vermont Avenue", "Connecticut Avenue", "Jail", "St. Charles Place", "Electric Company", "States Avenue", "Virginia Avenue","Pennsylvania Railroad", "St. James Place", "Community Chest","Tennessee Avenue","New York Avenue","Free Parking","Kentucky Avenue","Chance","Indiana Avenue","Illinois Avenue","B. & O. Railroad","Atlantic Avenue","Ventnor Avenue", "Water Works", "Marvin Gardens","Go To Jail","Pacific Avenue","North Carolina Avenue", "Community Chest", "Pennsylvania Avenue","Short Line", "Chance","Park Place", "Luxury Tax", "Boardwalk")
prices <- c("0","60","0","60","0","200","100","0","100","120","0","140","150","140","160","200","180","0","180","200","0","220","0","220","240","200","260","260","150","280","0","300","300","0","320","200","0","350","0","400")
prices <- as.numeric(prices)
#For rent that depends on 4 times the dice roll we preset the dice roll to be 7 as that is the most common 2 sided dice roll.
rent <- c("0","2","0","4","0","25","6","0","6","8","0","10","28","10","12","25","14","0","14","16","0","18","0","18","20","25","22","22","28","24","0","26","26","0","28","28","0","35","0","50")
rent <- as.numeric(rent)
ownerID<-rep(c(0),times=40)
board <- data.frame(numeric_tile = 1:40, tiles, prices,rent,ownerID)
(board)

#creating deck of chance cards
chance <- c("Advance to go (collect $200).", "Advance to Illinois Ave - If you pass Go, collect $200.", "Advance to St. Charles Place - If you pass Go, collect $200.", "Advance to nearest Utility. If unowned, you may buy it from the Bank. If owned, throw dice and pay owner a total ten times the amount thrown.", "Advance token to the nearest Railroad and pay owner twice the rental to which he/she is otherwise entitled. If Railroad is unowned, you may but it from the Bank.", "Bank pays you dividend of $50.", "Get Out of Jail Free.", "Go Back 3 Spaces.", "Go to Jail-Go directly to Jail-Do not pass Go, do not collect $200.", "Make general repairs on all your property-For each house pay $25-For each hotel $100.","Pay poor tax of $15","Take a trip to Reading Railroad–If you pass Go, collect $200.","Take a walk on the Boardwalk–Advance token to Boardwalk","You have been elected Chairman of the Board–Pay each player $50","Your building and loan matures—Collect $150","You have won a crossword competition—Collect $100")

chancecards <- data.frame(card_number = 1:16, chance)

#creating deck of community chest. Removed get out of jail free card and hotel/ house assessment cards
chest <- c("Advance to Go (Collect $200).","Bank error in your favor—Collect $200.","Doctor's fee—Pay $50.","From sale of stock you get $50.","Go to Jail–Go directly to jail–Do not pass Go–Do not collect $200.","Grand Opera Night—Collect $50 from every player for opening night seats.","Holiday Fund matures—Receive $100.","Income tax refund–Collect $20.","Life insurance matures–Collect $100.","Pay hospital fees of $100.","Pay school fees of $150.","Receive from services $25.","You have won second prize in a beauty contest–Collect $10.","You inherit $100.")

chestcards <- data.frame(card_number = 1:14, chest)

winner<-0 #this will be the global variable that stores the game winner's player ID

position<-1
money<-1000
lastRollSum<-0 #will need if player lands on utility (they owe 4x amount of dice)
ID<-0
turn<-0 #this will keep track of how many turns each player got to take
p1<- data.frame(ID,turn,position, money,lastRollSum)
p2<- data.frame(ID,turn,position, money, lastRollSum)
p3<- data.frame(ID,turn,position, money,lastRollSum)
p4<- data.frame(ID,turn,position, money,lastRollSum)
p1$ID<-1
p2$ID<-2
p3$ID<-3
p4$ID<-4

#for keeping track of locations. player vectors look at tile number and location logs look at tile name
p1vector = as.vector(1)
p2vector = as.vector(1)
p3vector = as.vector(1)
p4vector = as.vector(1)
loclogp1 = as.vector(board$tiles[1])
loclogp2 = as.vector(board$tiles[1])
loclogp3 = as.vector(board$tiles[1])
loclogp4 = as.vector(board$tiles[1])


getWinner<-function() #once we know the game has a winner, this function determines who that is and sets the global variable, winner, equal to the winner's player ID
  {
  if(p1$money!=0)
    {
    winner<<-1
    }
  if(p2$money!=0)
    {
    winner<<-2
    }
  if(p3$money!=0)
    {
    winner<<-3
    }
  if(p4$money!=0)
    {
    winner<<-4
    }
  }
isGameOver<-function() #after each turn, check is we have a winner (only one player left with money)
  {
  hasmoney<-0
  if(p1$money>0)
    {
    hasmoney<-hasmoney+1
    }
  if(p2$money>0)
    {
    hasmoney<-hasmoney+1
    }
  if(p3$money>0)
    {
    hasmoney<-hasmoney+1
    }
  if(p4$money>0)
    {
    hasmoney<-hasmoney+1
    }
  if(hasmoney==1) #if has money equals one, only one person has money left, therefore, they are the winner
    {
    getWinner() 
    return(TRUE)
    }
  return(FALSE)
  }
payPlayer<-function(id,amount) #checks who the owner of the property is and pays them
  {
  if(id==1)
    {
    p1$money<<-p1$money+amount
    }
  if(id==2)
    {
    p2$money<<-p2$money+amount
    }
  if(id==3)
    {
    p3$money<<-p3$money+amount
    }
  if(id==4)
    {
    p4$money<<-p4$money+amount
    }
  }
transferMoney<-function(p,to,amount) #updates player money
  {
  if(p$money < amount) #if player doesn't have enough to pay full price, pay what they have
    {
    amount<-p$money
    }
  p$money<-p$money-amount #subtracting the amount from the player who is paying
  payPlayer(to,amount)
  return(p)
  }
getOwner<-function(board,pos) #tells who property owner is
  {
  return(board$ownerID[pos])
} 
drawChestCard <- function(p) #contains all community chest card's and actions for each card
  {
    x <- sample(chest,1, replace = TRUE)
    
    
    if(x == "Advance to go (collect $50).") {
      p$position <- 1
      p$money <- p$money + 50
      
    }
    if(x =="Bank error in your favor—Collect $50.") {
      p$money <- p$money + 50
      
    }
    if(x =="Doctor's fee—Pay $50.") {
      p$money <- p$money - 50
      
    }
    if(x == "From sale of stock you get $50.") {
      p$money <- p$money +50
      
    }
    if(x == "Go to Jail–Go directly to jail–Do not pass Go–Do not collect $50.") {
      p$position <- 11
      p$money <- p$money - 50
      
    }
    if(x=="Grand Opera Night—Collect $50 from every player for opening night seats.") {
      p1$money <- p1$money - 50
      p2$money <- p2$money - 50
      p3$money <- p2$money - 50
      p4$money <- p4$money - 50
     
      p$money <- p$money + 200
    }
    if(x == "Holiday Fund matures—Receive $100.") {
      p$money <- p$money + 100
    }
    if(x == "Income tax refund–Collect $20.") {
      p$money <- p$money + 20
    }
    if(x == "Life insurance matures–Collect $100.") {
      p$money <- p$money + 100
    }
    if(x == "Pay hospital fees of $100.") {
      p$money <- p$money - 100
    }
    if(x == "Pay school fees of $150.") {
      p$money <- p$money - 150
    }
    if(x == "Receive from services $25.") {
      p$money <- p$money + 25
    }
    if(x == "You have won second prize in a beauty contest–Collect $10.") {
      p$money <- p$money + 10
    }
    if(x == "You inherit $100.") {
      p$money <- p$money + 100
    }
  return(p)
}
drawChanceCard <- function(p) #contains all chance card's and actions for each card
  {
    x <- sample(chestcards, 1, replace = TRUE)
    pos <- p$position
  if(x == "You have won a crossword competition—Collect $100") {
      p$money <- p$money + 100
    }
  if(x == "Your building and loan matures—Collect $150") {
      p$money <- p$money + 150
    }
  if(x == "Bank pays you dividend of $50.") {
      p$money <- p$money + 50
    }
  if(x=="Go to Jail-Go directly to Jail-Do not pass Go, do not collect $200.") {
      p$position <- 11
      p$money <- p$money - 50
    }
  if(x == "Go Back 3 Spaces."){
    p$position <- p$position - 3
  }
  if(x == "Pay poor tax of $15"){
    p$money <- p$money - 15
  }
  if(x =="Take a walk on the Boardwalk–Advance token to Boardwalk"){
    p$position <-40
  }
  if( x =="You have been elected Chairman of the Board–Pay each player $50"){
    p$money <- p$money - 200
    p1$money <- p1$money + 50
    p2$money <- p2$money + 50
    p3$money <- p3$money + 50
    p4$money <- p4$money + 50
  }
  if(x == "Advance to go (collect $200).") {
      p$position <- 1
      p$money <- p$money + 50
    }
  
  if(x == "Advance to St. Charles Place - If you pass Go, collect $200.") {
     p$position <- 12
    if(pos > p$position){
		#player passed go
		p$money<-p$money + 50
    }	
  }
  if( x == "Advance to Illinois Ave - If you pass Go, collect $200.") {
      p$position <- 25
      if(pos > p$position){
		#player passed go
		  p$money<-p$money + 50
       }	
    }
  if(x == "Advance to nearest Utility. If unowned, you may buy it from the Bank. If owned, throw dice and pay owner a total ten times the amount thrown.") {
    if(pos > 13 ) {
      p$position <- 29
    }
    else if(pos > 29) {
      p$position <- 29
    }
    else{
      p$position <- 13
    }
  }
    if(x == "Advance token to the #nearest Railroad and pay owner twice the rental to which he/she is otherwise entitled. If Railroad is unowned, you may but it from the Bank.") {
      if(p$position == 8) {
        p$position <- 6
      }
      if(p$position == 23) {
        p$position <- 26
      }
      if(p$position == 37) {
        p$position <- 6
        p$money <- p$money + 200
      }
    }
    if(x == "Take a trip to Reading Railroad–If you pass Go, collect $200.") {
     p$position <- 6
      if(pos > 6){
		#player passed go
		  p$money<-p$money + 50
       }	 
    }
  return(p)
}
addToPlayerVector<-function(p,pos) #checks what player is up and then updates location vectors
  {
    if(p$ID==1)
      {
      p1vector<<-append(p1vector,pos) #updates the player vector list of locations using tile number
      loclogp1<<-append(loclogp1,tiles[pos]) #updates the player vector list of locations using tile name
    }
    if(p$ID==2)
      {
      p2vector<<-append(p2vector,pos)
      loclogp2<<-append(loclogp2,tiles[pos])
      }
    if(p$ID==3)
      {
      p3vector<<-append(p3vector,p$position)
      loclogp3<<-append(loclogp3,tiles[pos])
      }
   if(p$ID==4)
      {
      p4vector<<-append(p4vector,p$position)
      loclogp4<<-append(loclogp4,tiles[pos])
      }
}
isLuxuryTax<-function(p,pos) #function checks to see if the player has landed on the luxury tax space
  {
  return(pos==39)
  }
isIncomeTax<-function(p,pos) #function checks to see if the player has landed on the income tax space
  {
  return(pos==5)
  }
isFreeParking<-function(pos) #function checks to see if the player has landed on free parking
  {
  return(pos==21)
  }
isCommunityChest<-function(pos) #function checks to see if the player has landed on community chest
  {
    return(pos==3|pos==18|pos==34) 
  }
isChance<-function(pos) #function checks to see if the player has landed on chance
  {
   return(pos==8|pos==23|pos==37) 
  }
isGoToJail<-function(p,pos) #function checks to see if the player has landed on the go to jail tile
  {
    return(pos==31)
  }
isJail<-function(pos) #function checks to see if the player has landed on the jail tile
  {
    return(pos==11)
  }
isGo<-function(pos) #function checks to see if the player has landed on Go
  {
    return(pos==1)
  }
isUtility<-function(pos) #function checks to see if player is on a utility, water works or electric company
  {
    return(pos==13|pos==29)
  }
isProperty<-function(pos) #function checks to see if the player is on a buyable property, not including utilities
  {
  return(board$prices[pos]>0 && pos!=13 && pos!=29) 
  } 
takeAction<-function(p) #function determines what action to take based on the player's location using if statements
  {
    if(p$position==0)
      {
      p$position<-p$position+1
      }
    if(isProperty(p$position)) 
      {
      ownerid<-getOwner(board,p$position) #checks if someone owns the property by looking at the ownerID column of the board data frame
      if(p$ID==ownerid) #this means the player already owns the property and their turn is over
        {
        return(p)
        }
      if(ownerid==0) #no one owns the property
        {
        if(p$money>=board$prices[p$position]) #check to see if player has enough money to buy property
          {
          p$money<-p$money-board$prices[p$position] #pay for property
          board$ownerID[p$position]<<-p$ID #assign ownership of property in board data frame
          return(p)
          }
        else
          {
          return(p) #if player doesn't have enough money to buy the property their turn is over
          }
        }
      else
        {
        if(isDone(p)) #if the player is out of the game we won't pay them
          {
          return(p)
          }
          else
            {
            p<-transferMoney(p, ownerid, board$rent[p$position])
            return(p)
            }
         }
      }
    if(isUtility(p$position))
      {
      ownerid<-getOwner(board,p$position) #checks if someone owns the property by looking at the ownerID column of the board data frame
      if(p$ID==ownerid) #this means the player already owns the property and their turn is over
        {
        return(p)
        }
      if(ownerid==0) #no one owns the property
        {
        if(p$money>=board$prices[p$position]) #check to see if player has enough money to buy property
          {
          p$money<-p$money-board$prices[p$position] #pay for property
          board$ownerID[p$position]<<-p$ID #assign ownership of property in board data frame
          return(p)
          }
        else
          {
          return(p) #if player doesn't have enough money to buy the property their turn is over
          }
        }
        else
        {
        if(isDone(p)) #if the player is out of the game we won't pay them
          {
          return(p)
          }
          else
            {
            amount<-p$lastRollSum *4
            p<-transferMoney(p, ownerid, amount) #rent is 4  times dice sum
            return(p)
            }
         }
      }
    if(isGoToJail(p,p$position))
      {
      p$money<-p$money-50
      p$position<-11
      return(p)
      }
    if(isJail(p$position)) #turn is over if player lands on jail space
      {
      return(p)
      }
    if(isGo(p$position)) #turn is over if player lands on Go
      {
      return(p)
      }
    if(isFreeParking(p$position)) #turn is over if player lands on free parking
      {
      return(p)
      }
    if(isCommunityChest(p$position))
      {
      p<-drawChestCard(p)
      }
  if(isIncomeTax(p,p$position))
    {
    p$money<-p$money-200 #pays $200 tax
    return(p)
    }
  if(isLuxuryTax(p,p$position))
    {
    p$money<-p$money-75 #pays $75 tax
    return(p)
    }
  if(isChance(p$position))
      {
      p<-drawChanceCard(p)
      }
    return(p)
  }
move<-function(p, pos) #function that updates the position of the player
  {
	if(pos < p$position)
	  {
		p$money<-p$money + 50 #player passed go
    }	
	p$position<-pos
	p<-takeAction(p)
	return(p)
}
getDesination<-function(pos, dicesum) #function that gets the new board position from adding the sum of the dice to current player position
    { 
    return((pos + dicesum) %% 40)
    }
roll<-function() #function to roll dice just using sampling 
    {
    return(sample(1:6,1))
    }
isDone<-function(p) #function checks if player taking a turn is out of money or not
    { 
	  return(p$money <= 0)
    }
takeTurn<-function(p)#function that includes all other functions that one player will go through to complete a turn
{
  if(isDone(p)){
    return(p)
  }
    p$turn<-p$turn+1 #the player has enough money to take a turn, so turn column increases by 1
    isTurnDone <- FALSE
    rollCount <- 0 #keeps track of how many rolls
    while(!isTurnDone) #loops to roll dice until thet do not get doubles
      { 
       rollCount<-rollCount+1
       d1<-roll()
       d2<-roll()
       p$lastRollSum<-d1+d2
       if(d1==d2 & rollCount==3) #this means the player has rolled doubles 3 times and is going to jail
         {
         p$money<-p$money-50 #to get out of jail player pays a $50 fee
         p$position<-11 #player goes to jail, tile 11
         return(p) #player's turn is over
       }
       if(d1!=d2) #if if gets here, the player did not roll doubles and the loop will end after getDestination is called
         {
         isTurnDone = TRUE
       }
      newPos<-getDesination(p$position, dicesum<-d1+d2)
      addToPlayerVector(p,newPos)
      p<-move(p, newPos)
    } 
  return(p)
}

while(isGameOver()==FALSE) #loops player's taking a turn until their is a winner
  {
p1<-takeTurn(p1)
isGameOver() #checks for a winner after each player's turn
p2<-takeTurn(p2)
isGameOver()
p3<-takeTurn(p3)
isGameOver()
p4<-takeTurn(p4)
isGameOver()
}
#to check the winner, print "winner" variable


```

